Testing cosmic isotropy and homogeneity with \textsl{Gaia} quasars

### 2023-06-17 paper/main.tex

% ## to-do items
% --------------
% - finish writing first draft
% - put in all references
% - address all XXX, YYY, ZZZ, HOGG, KSF in the text

@@

% text definitions
\shortauthors{hogg and storey-fisher}
\shorttitle{testing isotropy and homogeneity}
\newcommand{\documentname}{\textsl{Article}}
\newcommand{\sectionname}{Section}
\newcommand{\secref}[1]{\sectionname~\ref{#1}}
\newcommand{\figref}[1]{\figurename~\ref{#1}}

@@

\begin{abstract}\noindent
All cosmological models in general relativity are built on the assumption of large-scale isotropy and homogeneity.
These assumptions are testable.
Tests of cosmic isotropy and homogeneity are also strong tests of the quality of data or catalogs and their calibration.
Test precision increases as observations cover more solid angle, become more sensitive to distant objects, and are made with more uniform data.
The ESA \textsl{Gaia} Mission is primarily concerned with stars in the Milky Way, but it also observes millions of quasars, all sky (modulo Galactic extinction), to redshifts of around $4.5$.
It has produced the largest-comoving-volume public quasar catalog to date.
Here we use a carefully curated subsample of ZZZ \textsl{Gaia} quasars to show that the angular distribution, redshift distribution as a function of sky position, amplitude of clustering as a function of sky position, and counts of pairs as a function of separation (fractal dimension), are all consistent with large-scale isotropy and homogeneity at the precision possible given the size of the catalog and the amplitude of the large-scale structure.
Roughly speaking, the results can be summarized by saying that they show the Universe to be isotropic to better than XXX percent and homogeneous to better than YYY percent.
There is no sign of any North--South power asymmetry.
The isotropy results can also be interpreted as a strong test of the spectrophotometric uniformity of the \textsl{Gaia} Catalog.
\end{abstract}

\keywords{%
calibration
---
catalogs
---
cosmic~isotropy
---
cosmological~principle
---
cosmology
---
large-scale~structure~of~the~Universe
---
quasars
}
\section*{}
\clearpage
\section{Introduction}\label{sec:intro}

One of the fundamental assumptions of the physical cosmological model is that the Universe is isotropic and homogeneous on large scales (CITE).
This assumption is testable.
It is worthy of test, both because it is always worth testing fundamental assumptions, and because the large-scale structure has self-similar or fractal characteristics (CITE), which are characteristic of inhomogeneous distributions.
And, even in the ascendent physical model of cosmology, there is expected to be at least some power in the initial conditions on all length scale (CITE).

@@

All that said, we know now that the Universe is not a fractal on large scales (CITE).
There are other reasons to continue to test isotropy and homogeneity however.
One is that the cosmic microwave background appears to show a power asymmetry, in which there appears to be a larger amplitude of fluctuations in the North/South Galactic Cap than there is in the South/North (CITE).
If this asymmetry is real, it should show itself in the large-scale structure as well (CITE).

Another reason to make new tests of isotropy and homogeneity is that data samples keep getting larger in volume and number, permitting more and more sensitive tests.
Here we make use of the brand-new ESA \textsl{Gaia} DR3 data (CITE), which includes hundreds of thousands or even millions of quasars, all-sky (CITE).
This data set is not the largest cosmological data set in terms of number of tracers, but it is the largest-ever data set in terms of comoving volume, spanning some XXX $h^{-3}\,\Gpc^3$ (comoving), depending on how you want to define the sample volume.

A final but equally important reason to test isotropy and homogeneity is that these tests provide a very strong validation of the quality and uniformity of the data.
Indeed, the first work on homogeneity with the Sloan Digital Sky Survey data (CITE) was performed to test the calibration and stability of the data in preparation for the measurement of the baryon acoustic feature (CITE).
The work presented here represents a first strong test of the calibration and isotropy of the \textsl{Gaia} quasar sample.

\section{Data}\label{sec:data}

Hogg say: Make sure we cite the CU5 papers here and anything that helped make the extragalactic and quasar samples.

@@

\section{Isotropy}\label{sec:iso}

A simple kind of isotropy is evident in \figref{fig:lb}.
Because the sky plots shown there are equal-area projections, the uniformity of the sample in those plots is an indicator of isotropy.
This observation can be made more quantitative by projecting onto spherical harmonics.
In detail, we project onto the spherical harmonics by
\begin{equation}\label{eq:alms}
    a_{\ell m} = \sum_n Y_{\ell m}(\Omega_n) ~,
\end{equation}
where $a_{\ell m}$ is a spherical-harmonic amplitude,
$\ell, m$ are the angular degree and order,
$n$ is an integer index and the sum is over all quasars,
$Y_{\ell m}(\Omega)$ is the spherical harmonic at degree $\ell$ and order $m$,
and
$\Omega_n$ is the sky position (Galactic $l, b$) for quasar $n$.
The spherical harmonics are oriented in the natural way in the Galactic coordinate system, such that $Y_{10}$ has a maximum in the direction of the Galactic Center.
The top panel of \figref{fig:alms} shows these amplitudes $a_{\ell m}$.
The amplitudes $a_{\ell m}$ are complex, so \figref{fig:alms} shows the real and imaginary parts of the $m\geq 0$ orders.

@@

\begin{figure}[t!]
  \begin{mdframed}
  \color{captiongray}
  \begin{center}
    \includegraphics[width=\figurewidth]{notebooks/zhist.png}
  \end{center}
    \caption{The redshift distributions of the NGC and SGC subsamples. This demonstrates that the distribution of quasar spectrophotometric properties is very similar between the North and South.\label{fig:zhist}}
  \end{mdframed}
\end{figure}
HOGG: Another strong test of isotropy can be obtained from the redshift distribution.
Why is this strong?
It also connects to the hemispherical power asymmetry in the CMB.

@@

\section{Homogeneity}\label{sec:homo}

The weakest possible prediction of homogeneity is that there should be a mean density to the Universe.
This prediction is weak, but it has the advantage that it carries few other assumptions.
The prediction that there exists a mean density can be cast in terms of the scaling of pair counts or a measurement of the \emph{fractal dimension}.
The fractal dimension, for a set of points in a space, is defined to be the scaling of the total number of pairs $N_{DD}$ in the set with the maximum separation $L$ out to which pairs are counted.
If $N_{DD}$ is proportional to $L^d$ then the fractal dimension is $d$; if $d=3$ then the Universe has a mean density.

\begin{figure}[t!]
  \begin{mdframed}
  \color{captiongray}
  \begin{center}
    \includegraphics[width=\figurewidth]{notebooks/cumulativeDD_DR.png}
  \end{center}
    \caption{\textsl{Top:} The cumulative total number of quasar--quasar pairs as a function of separation for the NGC and SGC subsamples. Jackknife uncertainty estimates are shown but are smaller than the points for all points.
    \textsl{Bottom:} The same, but divided by the expectation in a homogeneous universe.
    In this case the expectation is estimated with $DR + RD - RR$ (see text for explanation).
    This figure demonstrates that the fractal dimension is extremely close to $3$.\label{fig:cumulative}}
  \end{mdframed}
\end{figure}
The cumulative number $N_{DD}$ of pairs up to separation $L$ is shown as a function of $L$ in the top panel of \figref{fig:cumulative}.
The top panel shows the raw counts of pairs, and the bottom panel shows the pairs relative to an expectation derived from the random catalog, which (by construction) has fractal dimension $d=3$.
Although the scaling in the top panel of \figref{fig:cumulative} looks very good, it can be biased by the finite volume within which we have observations; hence the ratios plotted in the bottom panel of \figref{fig:cumulative}.
In detail, the statistics that are plotted are cumulative pair counts obtained as follows:
We start by counting pairs within the quasar sample, between the quasar sample and the random catalog, and within the random catalog, by
\begin{align}\label{eq:DD}
    N_{DD}(L) &= \sum_n \sum_{n'\ne n} I(r_{nn'} < L) \\
    N_{DR}(L) &= f\,\sum_n \sum_{m'} I(r_{nm'} < L) \\
    N_{RD}(L) &= f\,\sum_m \sum_{n'} I(r_{mn'} < L) \\
    N_{RR}(L) &= f^2\,\sum_m \sum_{m'\ne m} I(r_{mm'} < L) ~,
\end{align}
where the sums over $n$ or $n'$ are over quasars,
the sums over $m$ or $m'$ are over random points,
$r_{nn'}$ is the comoving distance between quasar $n$ and quasar $n'$,
$I(r<L)$ is an indicator function that returns 1 if $r<L$ and 0 otherwise,
and $f$ is a dimensionless ratio of the density of the quasar sample (on the sky) to the density of the random catalog (on the sky).
With these counts in hand, the expectation $\hat{N}_{DD}(L)$ we use for the cumulative pair counts $N_{DD}(L)$ is given by
\begin{equation}
    \hat{N}_{DD}(L) = N_{DR}(L) + N_{RD}(L) - N_{RR}(L) ~,
\end{equation}
which is inspired by the currently standard correlation-function estimator (\citealt{ls}).

@@

For the maximum separation $L$ and the separations $r_{nn'}$ in \eqref{eq:DD} and the subsequent equations for the pair counts, we used the comoving distance in the \textsl{Planck} 2018 cosmology (\citealt{planck}).
This is a strong assumption, and wrong if the Universe has no mean density.
Indeed, since (as we note in \secref{sec:intro}) there are no known solutions to the equations of general relativity in an inhomogeneous universe, there is no way to compute pair separation distances $r_{nn'}$ without making the assumption of homogeneity.
The fact that this test passes---that the Universe appears to have a well-defined mean density and looks homogeneous on large scales---is a non-trivial test, but it does have this limitation, which cannot be addressed at present.

To put error bars on things... HOGG JACKKNIFE

\section{Statistical isotropy}\label{sec:stat}

If the fractal dimension is 3, then the correlation function should go to zero at large scales.
It does!
But the power asymmetry between the hemispheres challenges \emph{statistical isotropy} which is yet another test.

\begin{figure}[t!]
  \begin{mdframed}
  \color{captiongray}
  \begin{center}
    \includegraphics[width=\figurewidth]{notebooks/corrfunc.png}
  \end{center}
    \caption{The auto-correlation of the quasars as a function of separation for the NGC and SGC subsamples.
    This figure demonstrates that the clustering amplitude is similar between the North and the South.\label{fig:corrfunc}}
  \end{mdframed}
\end{figure}

\section{Discussion}\label{sec:discuss}

We found what we expected to find, but at very good precision.

How to summarize these results?

ESA \textsl{Gaia} rocks!

\begin{acknowledgements}
Be sure to cite all software: numpy, scipy, matplotlib, astropy.

Be sure to thank the Gaia~F\^ete and Gaia~Hike.

Be sure to put in the full Gaia acknowlegement.

Did we use 2MASS? If so, fix the text to capture it above and below.
\end{acknowledgements}

@@

\begin{thebibliography}{dummy}
\bibitem[Hoffleit \& Jaschek(1991)]{bsc}
Hoffleit, D. \& Jaschek, C.\ 1991, The Bright Star Catalog. New Haven: Yale University Observatory, 5th rev. ed.
\bibitem[Landy \& Szalay(1993)]{ls} Landy, S.~D. \& Szalay, A.~S.\ 1993, \apj, 412, 64. doi:10.1086/172900
\bibitem[Schlegel et al.(1998)]{sfd} Schlegel, D.~J., Finkbeiner, D.~P., \& Davis, M.\ 1998, \apj, 500, 525. doi:10.1086/305772
\end{thebibliography}

>>>>

% ## to-do items
% --------------
% - test the cmb dipole; cite Secrest et al. NO -- ABBY's PAPER -- BUT DISCUSS THIS POINT.
% - color choices?
% - explain, justify, or change how DD, DR, RD, and RR are counted -- the asymmetry between the tree and the sample compared to the tree.
% - discuss more philosophically why alms are sensitive tests of anisotropy
% - actually measure the fractal dimension
% - make some figure that quantitatively checks Gaia calibration
% - put the expectation hatN_DD onto the non-divided cumulative number plot
% - put in all missing references
% - address all CITE, XXX, YYY, ZZZ, HOGG, KSF in the text
% - make sure the figure axis labels and figure captions use precisely the math symbols used in the text; maybe also turn on TeX in matplotlib?

@@

% text definitions
\shortauthors{hogg and storey-fisher}
\shorttitle{testing isotropy and homogeneity}
\newcommand{\documentname}{\textsl{Article}}
\newcommand{\sectionname}{Section}
\newcommand{\secref}[1]{\sectionname~\ref{#1}}
\newcommand{\figref}[1]{\figurename~\ref{#1}}
\newcommand{\abby}[1]{\textbf{Abby says: #1}}

@@

\begin{abstract}\noindent
All cosmological models in general relativity are built on the assumption of large-scale isotropy and homogeneity.
These assumptions are testable.
Tests of cosmic isotropy and homogeneity are also strong tests of the quality of data or catalogs and their calibration.
Test precision increases as observations cover more solid angle, become more sensitive to distant objects, and are made with more uniform data.
The ESA \textsl{Gaia} Mission is primarily concerned with stars in the Milky Way, but it also observes millions of quasars, all sky (modulo Galactic extinction), to redshifts of around $4$.
It has produced the largest-comoving-volume public quasar catalog to date.
Here we use a carefully curated subsample of ZZZ \textsl{Gaia} quasars to show that the angular distribution, redshift distribution as a function of sky position, amplitude of clustering as a function of sky position, and counts of pairs as a function of separation (fractal dimension), are all consistent with large-scale isotropy and homogeneity at the precision possible given the size of the catalog and the amplitude of the large-scale structure.
Roughly speaking, the results can be summarized by saying that they show the Universe to be isotropic to better than XXX percent and homogeneous to better than YYY percent.
There is no sign of any North--South power asymmetry.
\abby{Is this the same hemispherical power asymmetry claimed from the CMB?}
The isotropy results can also be interpreted as a strong test of the spectrophotometric uniformity of the \textsl{Gaia} Catalog.
\end{abstract}

\keywords{%
calibration
---
catalogs
---
cosmic~isotropy
---
cosmological~principle
---
cosmology
---
large-scale~structure~of~the~Universe
---
quasars
}
\section*{}
\clearpage
\section{Introduction}\label{sec:intro}\noindent
One of the fundamental assumptions of the physical cosmological model is that the Universe is isotropic and homogeneous on large scales (CITE).
This assumption is testable.
It is worthy of test, both because it is always worth testing fundamental assumptions, and because the large-scale structure has self-similar or fractal characteristics (CITE), which are characteristic of inhomogeneous distributions.
And, even in the ascendent physical model of cosmology, there is expected to be at least some power in the initial conditions on all length scale (CITE).

@@

All that said, we know now that the Universe is not a fractal on large scales (\citealt{hogg05}; CITE OTHERS).
There are other reasons to continue to test isotropy and homogeneity however.
One is that the cosmic microwave background appears to show a power asymmetry, in which there appears to be a larger amplitude of fluctuations in the North/South Ecliptic (HOGG CHECK) Cap than there is in the South/North (\citealt{wmapanisotropy}; \citealt{planckanisotropy}).
If this asymmetry is real, it should show itself in the large-scale structure as well (\citealt{zhai}).

Another reason to make new tests of isotropy and homogeneity is that data samples keep getting larger in volume and number, permitting more and more sensitive tests.
Here we make use of the brand-new ESA \textsl{Gaia} DR3 data (CITE), which includes hundreds of thousands or even millions of quasars, all-sky (CITE).
This data set is not the largest cosmological data set in terms of number of tracers, but it is the largest-ever data set in terms of comoving volume, spanning some XXX $h^{-3}\,\Gpc^3$ (comoving), depending on how you want to define the sample volume.
Recent tests of homogeneity include ZZZ and ZZZ; the samples used in those studies were smaller by a factor of QQQ and WWW (approximately) in comoving volume, relative to the quasar catalog we use here.

A final but equally important reason to test isotropy and homogeneity is that these tests provide a very strong validation of the quality and uniformity of the data.
Indeed, the first work on homogeneity with the Sloan Digital Sky Survey data (\citealt{hogg05}) was performed to test the calibration and stability of the data in preparation for the measurement of the baryon acoustic feature (\citealt{baf}).
The work presented here represents a first strong test of the calibration of the \textsl{Gaia} quasar sample and associated photometry and spectroscopy; it has implications for the accuracy and uniformity of the \textsl{Gaia} data more generally.

\section{Data}\label{sec:data}\noindent
Hogg say: Make sure we cite the CU5 papers and XP papers here and anything that helped make the extragalactic and quasar samples.

@@

\section{Isotropy}\label{sec:iso}\noindent
A simple kind of isotropy is evident in \figref{fig:lb}:
Because the sky plots shown there are equal-area projections, the uniformity of the sample in those plots is an indicator of isotropy.
This observation can be made more quantitative by projecting onto spherical harmonics.
In detail, we project onto the spherical harmonics by
\begin{equation}\label{eq:alms}
    a_{\ell m} = \sum_n Y_{\ell m}(\Omega_n) ~,
\end{equation}
where $a_{\ell m}$ is a spherical-harmonic amplitude,
$\ell, m$ are the angular degree and order,
$n$ is an integer index and the sum is over all quasars,
$Y_{\ell m}(\Omega)$ is the spherical harmonic at degree $\ell$ and order $m$,
and
$\Omega_n$ is the sky position (Galactic $l, b$) for quasar $n$.
The spherical harmonics are oriented in the natural way in the Galactic coordinate system, such that $Y_{10}$ has a maximum in the direction of the Galactic Center.
The top panel of \figref{fig:alms} shows these amplitudes $a_{\ell m}$.
The amplitudes $a_{\ell m}$ are complex, so \figref{fig:alms} shows the real and imaginary parts of the $m\geq 0$ orders.

@@

\begin{figure}[t!]
  \begin{mdframed}
  \color{captiongray}
  \begin{center}
    \includegraphics[width=\figurewidth]{notebooks/zhist.png}
  \end{center}
    \caption{The redshift distributions of the NGC and SGC subsamples. This demonstrates that the distribution of quasar spectrophotometric properties is very similar between the North and South.\label{fig:zhist}}
  \end{mdframed}
\end{figure}
HOGG: Another strong test of isotropy can be obtained from the redshift distribution.
Why is this strong?
It also connects to the hemispherical power asymmetry in the CMB (\citealt{wmapanisotropy}; \citealt{planckanisotropy}).

@@

\section{Homogeneity}\label{sec:homo}\noindent
The weakest possible prediction of homogeneity is that there should be a mean density to the Universe.
This prediction is weak, but it has the advantage that it carries few other assumptions.
The prediction that there exists a mean density can be cast in terms of the scaling of pair counts or a measurement of the \emph{fractal dimension}.
The fractal dimension, for a set of points in a space, is defined to be the scaling of the total number of pairs $N_{DD}$ in the set with the maximum separation $s$ out to which pairs are counted.
If $N_{DD}$ is proportional to $s^d$ then the fractal dimension is $d$; if $d=3$ then the Universe has a mean density.

\begin{figure}[t!]
  \begin{mdframed}
  \color{captiongray}
  \begin{center}
    \includegraphics[width=\figurewidth]{notebooks/cumulativeDD_DR.png}
  \end{center}
    \caption{\textsl{Top:} The cumulative total number $N_{DD}$ \eqref{eq:DD} of quasar--quasar pairs as a function of comoving separation $s$ for the NGC and SGC subsamples. Jackknife uncertainty estimates are shown but are smaller than the points for all points.
    \textsl{Bottom:} The same, but divided by the expectation $\hat{N}_{DD}$ \eqref{eq:DDhat} in a homogeneous universe.
    This figure demonstrates that the fractal dimension is extremely close to $3$.\label{fig:cumulative}}
  \end{mdframed}
\end{figure}
The cumulative number $N_{DD}$ of pairs up to separation $s$ is shown as a function of $s$ in the top panel of \figref{fig:cumulative}.
The top panel shows the raw counts of pairs, and the bottom panel shows the pairs relative to an expectation derived from the random catalog, which (by construction) has fractal dimension $d=3$.
Although the scaling in the top panel of \figref{fig:cumulative} looks very good, it can be biased by the finite volume within which we have observations; hence the ratios plotted in the bottom panel of \figref{fig:cumulative}.
In detail, the statistics that are plotted are cumulative pair counts obtained as follows:
We start by counting pairs within the quasar sample, between the quasar sample and the random catalog, and within the random catalog, by
\begin{align}\label{eq:DD}
    N_{DD}(s) &= \sum_n \sum_{n'\ne n} I(r_{nn'} < s) \\
    N_{DR}(s) &= f\,\sum_n \sum_{m'} I(r_{nm'} < s) \\
    N_{RD}(s) &= f\,\sum_m \sum_{n'} I(r_{mn'} < s) \\
    N_{RR}(s) &= f^2\,\sum_m \sum_{m'\ne m} I(r_{mm'} < s)\label{eq:RR} ~,
\end{align}
where the sums over $n$ or $n'$ are over quasars,
the sums over $m$ or $m'$ are over random points,
$r_{nn'}$ is the comoving distance between quasar $n$ and quasar $n'$,
$I(r<s)$ is an indicator function that returns 1 if $r<s$ and 0 otherwise,
and $f$ is a dimensionless ratio of the density of the quasar sample (on the sky) to the density of the random catalog (on the sky).
With these counts in hand, the expectation $\hat{N}_{DD}(s)$ we use for the cumulative pair counts $N_{DD}(s)$ is given by
\begin{equation}\label{eq:DDhat}
    \hat{N}_{DD}(s) = N_{DR}(s) + N_{RD}(s) - N_{RR}(s) ~,
\end{equation}
which is inspired by the currently standard correlation-function estimator (\citealt{ls}).

@@

For the maximum separation $s$ and the separations $r_{nn'}$ in \eqref{eq:DD} and the subsequent equations for the pair counts, we used the comoving distance in the \textsl{Planck} 2018 cosmology (\citealt{planck18}).
This is a strong assumption, and wrong if the Universe has no mean density.
Indeed, since (as we note in \secref{sec:intro}) there are no known solutions to the equations of general relativity in an inhomogeneous universe, there is no way to compute pair separation distances $r_{nn'}$ without making the assumption of homogeneity.
The fact that this test passes---that the Universe appears to have a well-defined mean density and looks homogeneous on large scales---is a non-trivial test, but it does have this limitation, which cannot be addressed at present.

In order to put error bars on the counts $N_{DD}$ and its expectation $\hat{N}_{DD}$ (HOGG: DID WE DO THAT?), we performed jackknife resampling.
Because we want the jackknife resampling to estimate uncertainty due not just to shot noise in the sample, but also large-scale structure modes, we performed the jackknife with 12 subsamples, in each of which we left out a disjoint contiguous 1/12 of the sky.
In detail we segmented the sky by galactic longitude $l$, and in each of the 12 leave-1/12-out subsample we re-did the full analysis including all counts of pairs.
The error bars in \figref{fig:cumulative} are estimated by this jackknife.

\section{Statistical isotropy}\label{sec:stat}\noindent
If the fractal dimension is 3, then the correlation function should go to zero at large scales.
It does (CITE THINGS)!
But observations of the cosmic microwave background have suggested that there might be a power asymmetry between the North and South Galactic Caps (CITE).
This asymmetry claim constitutes a claim of statistical anisotropy:
The statistics of the matter distribution depend on direction.
In a homogeneous universe (homogeneous in the sense of having a mean density), that statistical anisotropy, in turn, would imply a statistical inhomogeneity.
Thus it is interesting to check the quasar sample for any evidence for such a statistical anisotropy or statistical inhomogeneity.

The pair counts $N_{DD}(s), N_{DR}(s), N_{RD}(s), N_{RR}(s)$ obtained in \eqref{eq:DD} through \eqref{eq:RR} can be massaged into differential counts (counts within a separation bin) and then used to construct an estimate of the 3-space auto-correlation function of the quasars.
The differential counts $\Delta N_{DD,i}, \Delta N_{DR,i}, \Delta N_{RD, i}, \Delta N_{RR, i}$ in a separation bin $i$ can be defined with differences like
\begin{equation}
    \Delta N_{DD,i} = N_{DD}(s_i+\frac{1}{2}\,\Delta s) - N_{DD}(s_i-\frac{1}{2}\Delta s) ~,
\end{equation}
where $s_i$ is the bin center and $\Delta s$ is the bin width in separation space.
These differential counts can be combined into an estimate $\hat\xi(s_i)$ of the auto-correlation function with the Landy--Szalay estimator (\citealt{ls}):
\begin{equation}\label{eq:ls}
    \hat\xi(s_i) = \frac{\Delta N_{DD,i} - \Delta N_{DR,i} - \Delta N_{RD,i} + \Delta N_{RR,i}}{\Delta N_{RR,i}} ~.
\end{equation}

\begin{figure}[t!]
  \begin{mdframed}
  \color{captiongray}
  \begin{center}
    \includegraphics[width=\figurewidth]{notebooks/corrfunc.png}
  \end{center}
    \caption{The auto-correlation of the quasars as a function of comoving separation $s$ for the NGC and SGC subsamples, as estimated according to \eqref{eq:ls}.
    This figure demonstrates that the clustering amplitude is similar between the North and the South.\label{fig:corrfunc}}
  \end{mdframed}
\end{figure}
\figref{fig:corrfunc} shows the quasar auto-correlation function estimated according to \eqref{eq:ls} for the quasar sample in coarse separation bins, separated by sky hemsiphere.
The figure shows that the auto-correlation function is statistically consistent between hemispheres for this sample.
This result does not confirm the hemispherical power asymmetry observed in the cosmic microwave background.
Instead it supports the conclusion that the Universe is statistically isotropic and statistically homogeneous.

\section{Discussion}\label{sec:discuss}\noindent
HOGG: We found what we expected to find, but at very good precision.

HOGG: How to summarize these results?

HOGG: ESA \textsl{Gaia} rocks!

\begin{acknowledgements}
It is a pleasure to thank
  Anna-Christina Eilers (MIT),
  Hans-Walter Rix (MPIA), and
  Abby Williams (NYU, Caltech) for valuable discussions. KSF: OTHERS?
This project was developed in part at the 2022 Gaia F\^ete, hosted by the Center for Computational Astrophysics at the Flatiron Institute.
This project was also developed in part at the Gaia Hike, a workshop hosted by the University of British Columbia and the Canadian Institute for Theoretical Astrophysics in 2022 June.

HOGG: Be sure to put in the full Gaia acknowlegement.

HOGG: Did we use 2MASS? If so, fix the text to capture it above and below.
\end{acknowledgements}

@@

\begin{thebibliography}{dummy}
\bibitem[Eisenstein et al.(2005)]{baf} Eisenstein, D.~J., Zehavi, I., Hogg, D.~W., et al.\ 2005, \apj, 633, 560. doi:10.1086/466512
\bibitem[Eriksen et al.(2004)]{wmapanisotropy} Eriksen, H.~K., Hansen, F.~K., Banday, A.~J., et al.\ 2004, \apj, 605, 14. doi:10.1086/382267
\bibitem[Hoffleit \& Jaschek(1991)]{bsc}
Hoffleit, D. \& Jaschek, C.\ 1991, The Bright Star Catalog. New Haven: Yale University Observatory, 5th rev. ed.
\bibitem[Hogg et al.(2005)]{hogg05} Hogg, D.~W., Eisenstein, D.~J., Blanton, M.~R., et al.\ 2005, \apj, 624, 54. doi:10.1086/429084
\bibitem[Landy \& Szalay(1993)]{ls} Landy, S.~D. \& Szalay, A.~S.\ 1993, \apj, 412, 64. doi:10.1086/172900
\bibitem[Planck Collaboration(2014)]{planckanisotropy} Planck Collaboration, Ade, P.~A.~R., Aghanim, N., et al.\ 2014, \aap, 571, A23. doi:10.1051/0004-6361/201321534
\bibitem[Planck Collaboration(2020)]{planck18} Planck Collaboration, Aghanim, N., Akrami, Y., et al.\ 2020, \aap, 641, A6. doi:10.1051/0004-6361/201833910
\bibitem[Schlegel et al.(1998)]{sfd} Schlegel, D.~J., Finkbeiner, D.~P., \& Davis, M.\ 1998, \apj, 500, 525. doi:10.1086/305772
\bibitem[Storey-Fisher et al.(2022)]{ksf} Storey-Fisher, K, Hogg, D.~W., Eilers, A.-C., et al.\ 2022, in preparation
\bibitem[Zhai \& Blanton(2017)]{zhai} Zhai, Z. \& Blanton, M.~R.\ 2017, \apj, 850, 41. doi:10.3847/1538-4357/aa93e1
\end{thebibliography}
